---
title: "Lab4_exploratory"
author: "Rosanna Neuhausler"
date: "11/10/2017"
output: pdf_document
---

```{r echo = FALSE, message=FALSE, warning=FALSE}
#install.packages('e1071')
library(dplyr)
library(ggplot2)
library(reshape2)
library(e1071)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#Theme setup for plotting
blank_theme <- theme_minimal() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(size = 10), 
        axis.title=element_text(size=10))
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
# Get the data for three images
path <- 'image_data/'
image1 <- read.table(paste0(path, 'image1.txt'), header = F)
image2 <- read.table(paste0(path, 'image2.txt'), header = F)
image3 <- read.table(paste0(path, 'image3.txt'), header = F)

# Add informative column names.
collabs <- c('y','x','label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
names(image1) <- collabs
names(image2) <- collabs
names(image3) <- collabs
```


## 1 EDA

1. Plot the expert labels for the presence or absence of clouds, according to a map (i.e. use the X, Y coordinates).

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height = 3, fig.align = "center"}
# Plot the expert pixel-level classification

image1$image = 1
image2$image = 2
image3$image = 3

images <- dplyr::bind_rows(image1, image2)
images <- dplyr::bind_rows(images, image3)

images %>%
  ggplot() + 
  geom_point(aes(x = x, y = y, color = factor(label))) + 
  facet_wrap(~ image, ncol = 3) +
  blank_theme +
  guides(color=guide_legend("label")) +
  theme(legend.position="bottom") +
  labs(title = "Expert Labels",
       caption = "Expert labels, -1, 0, 1 representing no cloud, unknown, and cloud, respectively", 
       x = "latitude", y = "longitude") 

```

2. Explore the relationships between the radiances of different angles, both visually and quantitatively.

Do you notice differences between the two classes (cloud, no cloud) based on the radiances? 

Are there differences based on the features (CORR, NDAI, SD)?


```{r echo = FALSE, message=FALSE, warning=FALSE}

panel.cor = function(x, y, digits=2, prefix="", cex.cor, ...){
  usr = par("usr")
  on.exit(par(usr))
  par(usr = c(0,1,0,1))
  r = abs(cor(x,y,use='complete.obs'))
  txt=format(c(r,0.123456789),digits=digits)[1]
  txt=paste(prefix,txt,sep='')
  if(missing(cex.cor)) cex.cor = 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor*(1+r)/2)
}

panel.hist = function(x, ...){
  usr = par('usr')
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5))
  h = hist(x, plot=FALSE)
  breaks = h$breaks
  nB = length(breaks)
  y=h$counts
  y=y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, ...)
}

```


2. Explore the relationships between the radiances of different angles, both visually and quantitatively.

There seems to be a trend in the difference between these angles.


```{r echo = FALSE, message=FALSE, warning=FALSE}

image1 %>%
  filter(label == -1) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 0) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 1) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:6, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable) +
  blank_theme +
  guides(color=guide_legend("radiance")) +
  theme(legend.position="bottom") +
  labs(title = "Angle images for image 1",
       caption = " ", 
       x = "latitude", y = "longitude") 


#generally shifted, maybe add a gaussian filter?
```


Do you notice differences between the two classes (cloud, no cloud) based on the radiances? Are there

```{r echo = FALSE, message=FALSE, warning=FALSE}

d <- melt(image1, id = 1:6, measure = 7:11)

p <- ggplot(d, aes(x = value, group = factor(label), fill = factor(label))) + 
  geom_density(alpha = 0.5) 
  scale_fill_discrete(name = "Expert label")
p + facet_wrap(~ variable, ncol = 3, scales = "free") + blank_theme
```

differences based on the features (CORR, NDAI, SD)?


```{r echo = FALSE, message=FALSE, warning=FALSE}

d <- melt(image1, id = 1:3, measure = 4:6)

p = ggplot(d, aes(x = value, group = factor(label), fill = factor(label))) + 
  geom_density(alpha = 0.5) 
  scale_fill_discrete(name = "Expert label")
p + facet_wrap(~ variable, ncol = 3, scales = "free" ) +
  blank_theme
```





```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image2, id = 1:6, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image3, id = 1:3, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable)
```



```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:3, measure = 4:6)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~variable, scales = "free")
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:3, measure = 4:6)


p <- ggplot(d,aes(value)) +
  geom_bar()
p+ facet_wrap(label ~ variable)
```






```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = NDAI)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = SD)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = CORR)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 0) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == -1) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 1) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = NDAI)) +
  geom_point() +
  facet_wrap(~ label)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = CORR)) +
  geom_point() +
  facet_wrap(~ label)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = SD)) +
  geom_point() +
  facet_wrap(~ label)
```


