---
title: "Lab4_exploratory"
author: "Rosanna Neuhausler"
date: "11/10/2017"
output: pdf_document
---

```{r echo = FALSE, message=FALSE, warning=FALSE}
#install.packages('e1071')

library(dplyr)
library(ggplot2)
library(reshape2)
library(e1071)


```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#Theme setup for plotting
blank_theme <- theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(size = 10), 
        axis.title=element_text(size=10))
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
# Get the data for three images
path <- 'image_data/'
image1 <- read.table(paste0(path, 'image1.txt'), header = F)
image2 <- read.table(paste0(path, 'image2.txt'), header = F)
image3 <- read.table(paste0(path, 'image3.txt'), header = F)

# Add informative column names.
collabs <- c('y','x','label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
names(image1) <- collabs
names(image2) <- collabs
names(image3) <- collabs
```


## 1 EDA


```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height = 3, fig.align = "center"}
# Plot the expert pixel-level classification

image1$image = 1
image2$image = 2
image3$image = 3

images <- dplyr::bind_rows(image1, image2)
images <- dplyr::bind_rows(images, image3)

images %>%
  ggplot() + 
  geom_point(aes(x = x, y = y, color = factor(label))) + 
  facet_wrap(~ image, ncol = 3) +
  ggtitle('') + 
  blank_theme
```

```{r echo = FALSE, message=FALSE, warning=FALSE}

panel.cor = function(x, y, digits=2, prefix="", cex.cor, ...){
  usr = par("usr")
  on.exit(par(usr))
  par(usr = c(0,1,0,1))
  r = abs(cor(x,y,use='complete.obs'))
  txt=format(c(r,0.123456789),digits=digits)[1]
  txt=paste(prefix,txt,sep='')
  if(missing(cex.cor)) cex.cor = 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor*(1+r)/2)
}

panel.hist = function(x, ...){
  usr = par('usr')
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5))
  h = hist(x, plot=FALSE)
  breaks = h$breaks
  nB = length(breaks)
  y=h$counts
  y=y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, ...)
}

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#image1$type = image1$label + 1
#cols <- character(nrow(image1))
#cols[] <- "black"
#cols[image1$label == 0] <- "blue"
#cols[image1$label == 1] <- "red"

image1 %>%
  filter(label == -1) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 0) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 1) %>%
  select('DF','CF','BF','AF','AN') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```



```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:6, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable)


#generally shifted, maybe add a gaussian filter?
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:6, measure = 7:11)

ggplot(d, aes(x, y, color = label)) +
  geom_point() +
  facet_wrap(~ variable)


#generally shifted, maybe add a gaussian filter?
```


Explore the relationships between the radiances of different angles, both visually and quan- titatively.
Do you notice differences between the two classes (cloud, no cloud) based on the radiances? Are there
differences based on the features (CORR, NDAI, SD)?



```{r echo = FALSE, message=FALSE, warning=FALSE}

d <- melt(image1, id = 1:6, measure = 7:11)

p <- ggplot(d, aes(x = value, group = factor(label), fill = factor(label))) + 
  geom_density(alpha = 0.5) 
  scale_fill_discrete(name = "Expert label")
p + facet_wrap(~ variable, ncol = 3 ) + blank_theme
```

```{r echo = FALSE, message=FALSE, warning=FALSE}

d <- melt(image1, id = 1:3, measure = 4:5)

p = ggplot(d, aes(x = value, group = factor(label), fill = factor(label))) + 
  geom_density(alpha = 0.5) 
  scale_fill_discrete(name = "Expert label")
p + facet_wrap(~ variable, ncol = 3 )
```





```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image2, id = 1:6, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image3, id = 1:3, measure = 7:11)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~ variable)
```



```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:3, measure = 4:6)

ggplot(d, aes(x, y, color = value)) +
  geom_point() +
  facet_wrap(~variable)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
d <- melt(image1, id = 1:3, measure = 4:6)


p <- ggplot(d,aes(value)) +
  geom_bar()
p+ facet_wrap(label ~ variable)
```






```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = NDAI)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = SD)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = CORR)) +
  geom_point()
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 0) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == -1) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```


```{r echo = FALSE, message=FALSE, warning=FALSE}
image1 %>%
  filter(label == 1) %>%
  select('NDAI','SD','CORR') %>%
  pairs(upper.panel = panel.cor,
        diag.panel = panel.hist,
        lower.panel = panel.smooth,
        cex.labels = 2, font.labels = 2)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = NDAI)) +
  geom_point() +
  facet_wrap(~ label)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = CORR)) +
  geom_point() +
  facet_wrap(~ label)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
ggplot(image1, aes(x, y, color = SD)) +
  geom_point() +
  facet_wrap(~ label)
```


Explore the relationships between the radiances of different angles, both visually and quan- titatively.
Do you notice differences between the two classes (cloud, no cloud) based on the radiances? Are there
differences based on the features (CORR, NDAI, SD)?

collabs <- c('y','x','label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
