---
title: "Exploring the random forest"
author: "Bryan, Nan"
date: "11/12/2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r load_data, cache = TRUE, echo = FALSE}
library(knitr)
library(tidyverse)
library(glmnet)
library(ranger)
library(dplyr)
# load data
image1 <- read.table('../image_data/image1.txt', header = F)
image2 <- read.table('../image_data/image2.txt', header = F)
image3 <- read.table('../image_data/image3.txt', header = F)

# add column labels
collabs <- c('y','x','label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
names(image1) <- collabs
names(image2) <- collabs
names(image3) <- collabs
```

```{r load_utils, cache = FALSE, echo = FALSE}
# TODO: put ths smoothing_fun into a common utilities folder
source('../LASSO/smoothing_fun.R')
source('../LASSO/lasso_lib.R')
```

# Fit using random forest
We fit a random forest to image 1, and examine the predictive accuracy on the second image. 

```{r get_features, cache = TRUE}
# get indices with labels
image1_labeled_indx <- which(image1$label != 0)
image2_labeled_indx <- which(image2$label != 0)
image3_labeled_indx <- which(image3$label != 0)

feature_names <- colnames(image1)[3:dim(image1)[2]]

image1_labeled <- image1[image1_labeled_indx, feature_names]
image2_labeled <- image2[image2_labeled_indx, feature_names]
image3_labeled <- image3[image3_labeled_indx, feature_names]

image1_labeled$label <- as.factor(image1_labeled$label)
image2_labeled$label <- as.factor(image2_labeled$label)
image3_labeled$label <- as.factor(image3_labeled$label)

rf_fit_orig <- ranger(label ~ ., data = image1_labeled, 
                 num.trees = 500, 
                 probability = TRUE, 
                 importance = 'impurity')
```

We print the predictive accuracy: 

```{r, cache = TRUE}
# predict on image1
image1_prediction_prob_rf <- predict(rf_fit_orig, 
                              dat = image1_labeled, 
                              type = 'response')

image1_prediction_rf <- get_labels_from_prob(image1_prediction_prob_rf$predictions[, 2])


# predict on image2
image2_prediction_prob_rf <- predict(rf_fit_orig, 
                              dat = image2_labeled, 
                              type = 'response')

image2_prediction_rf <- get_labels_from_prob(image2_prediction_prob_rf$predictions[, 2])

# predict on image3
image3_prediction_prob_rf <- predict(rf_fit_orig, 
                              dat = image3_labeled, 
                              type = 'response')

image3_prediction_rf <- get_labels_from_prob(image3_prediction_prob_rf$predictions[, 2])

cat('Predictive accuracy on the label pixels of image 1: \n')
cat('note that image 1 was used to fit the RF: \n')
print(mean(image1$label[image1_labeled_indx] == image1_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 2: \n')
print(mean(image2$label[image2_labeled_indx] == image2_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 3: \n')
print(mean(image3$label[image3_labeled_indx] == image3_prediction_rf))
```

And look at feature importances

```{r, cache = TRUE}
cat('feature importances by decrease in GINI coefficient: \n')
print(sort(importance(rf_fit_orig) / sum(importance(rf_fit_orig)), decreasing = T))
```


# Lets try to combine the first two images and test on the third

```{r, cache = TRUE}
rf_fit_two <- ranger(label ~ ., data = rbind(image1_labeled, image2_labeled), 
                 num.trees = 500, 
                 probability = TRUE, 
                 importance = 'impurity')
```

Print out the training and testing prediction accuracy

```{r, cache = TRUE}
# predict on image1
image1_prediction_prob_rf <- predict(rf_fit_two, 
                              dat = image1_labeled, 
                              type = 'response')

image1_prediction_rf <- get_labels_from_prob(image1_prediction_prob_rf$predictions[, 2])


# predict on image2
image2_prediction_prob_rf <- predict(rf_fit_two, 
                              dat = image2_labeled, 
                              type = 'response')

image2_prediction_rf <- get_labels_from_prob(image2_prediction_prob_rf$predictions[, 2])

# predict on image3
image3_prediction_prob_rf <- predict(rf_fit_two, 
                              dat = image3_labeled, 
                              type = 'response')

image3_prediction_rf <- get_labels_from_prob(image3_prediction_prob_rf$predictions[, 2])

cat('Predictive accuracy on the label pixels of image 1: \n')
cat('note that image 1 was used to fit the RF: \n')
print(mean(image1$label[image1_labeled_indx] == image1_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 2: \n')
print(mean(image2$label[image2_labeled_indx] == image2_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 3: \n')
print(mean(image3$label[image3_labeled_indx] == image3_prediction_rf))
```

# Image 3 seems a little bit different. If we train a model on 3 and predict on 1 and 2:

```{r}
rf_fit_3 <- ranger(label ~ ., data = image3_labeled, 
                 num.trees = 500, 
                 probability = TRUE, 
                 importance = 'impurity')

image1_prediction_prob_rf <- predict(rf_fit_3, 
                              dat = image1_labeled, 
                              type = 'response')
image1_prediction_rf <- get_labels_from_prob(image1_prediction_prob_rf$predictions[, 2])

# predict on image2
image2_prediction_prob_rf <- predict(rf_fit_3, 
                              dat = image2_labeled, 
                              type = 'response')
image2_prediction_rf <- get_labels_from_prob(image2_prediction_prob_rf$predictions[, 2])

# predict on image3
image3_prediction_prob_rf <- predict(rf_fit_3, 
                              dat = image3_labeled, 
                              type = 'response')
image3_prediction_rf <- get_labels_from_prob(image3_prediction_prob_rf$predictions[, 2])

cat('Predictive accuracy on the label pixels of image 1: \n')
print(mean(image1$label[image1_labeled_indx] == image1_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 2: \n')
print(mean(image2$label[image2_labeled_indx] == image2_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 3: \n')
cat('note that image 3 was used to fit the RF: \n')
print(mean(image3$label[image3_labeled_indx] == image3_prediction_rf))
```

# Try to add neighboring information and run random forest again

```{r}
add_neighbor_info <- function(image){
  image_l <- image %>%
    mutate(x = x - 1) %>% 
    select(-label)
  image_r <- image %>%
    mutate(x = x + 1) %>% 
    select(-label)
  image_u <- image %>%
    mutate(y = y + 1) %>% 
    select(-label)
  image_d <- image %>%
    mutate(y = y - 1) %>% 
    select(-label)
  image_neighbor <- merge(image, image_l, 
                          by = c('x', 'y'), suffixes = c('','_l'))
  image_neighbor <- merge(image_neighbor, image_r, 
                          by = c('x', 'y'), suffixes = c('','_r'))
  image_neighbor <- merge(image_neighbor, image_u, 
                          by = c('x', 'y'), suffixes = c('','_u'))
  image_neighbor <- merge(image_neighbor, image_d, 
                          by = c('x', 'y'), suffixes = c('','_d'))
  return(image_neighbor)
}
image1_neighbor <- add_neighbor_info(image1)
image2_neighbor <- add_neighbor_info(image2)
image3_neighbor <- add_neighbor_info(image3)

image1_neighbor_labeled_indx <- which(image1_neighbor$label != 0)
image2_neighbor_labeled_indx <- which(image2_neighbor$label != 0)
image3_neighbor_labeled_indx <- which(image3_neighbor$label != 0)

feature_names <- colnames(image1_neighbor)[3:dim(image1_neighbor)[2]]

image1_neighbor_labeled <- image1_neighbor[image1_neighbor_labeled_indx, feature_names]
image2_neighbor_labeled <- image2_neighbor[image2_neighbor_labeled_indx, feature_names]
image3_neighbor_labeled <- image3_neighbor[image3_neighbor_labeled_indx, feature_names]

image1_neighbor_labeled$label <- as.factor(image1_neighbor_labeled$label)
image2_neighbor_labeled$label <- as.factor(image2_neighbor_labeled$label)
image3_neighbor_labeled$label <- as.factor(image3_neighbor_labeled$label)

# fit model
rf_fit_neighbor <- ranger(label ~ ., data = image1_neighbor_labeled, 
                 num.trees = 500, 
                 probability = TRUE, 
                 importance = 'impurity')

# predict using the model
image1_prediction_prob_rf <- predict(rf_fit_neighbor, 
                              dat = image1_neighbor_labeled, 
                              type = 'response')

image1_prediction_rf <- get_labels_from_prob(image1_prediction_prob_rf$predictions[, 2])


# predict on image2
image2_prediction_prob_rf <- predict(rf_fit_neighbor, 
                              dat = image2_neighbor_labeled, 
                              type = 'response')

image2_prediction_rf <- get_labels_from_prob(image2_prediction_prob_rf$predictions[, 2])

# predict on image3
image3_prediction_prob_rf <- predict(rf_fit_neighbor, 
                              dat = image3_neighbor_labeled, 
                              type = 'response')

image3_prediction_rf <- get_labels_from_prob(image3_prediction_prob_rf$predictions[, 2])

cat('Predictive accuracy on the label pixels of image 1: \n')
cat('note that image 1 was used to fit the RF: \n')
print(mean(image1_neighbor$label[image1_neighbor_labeled_indx] == image1_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 2: \n')
print(mean(image2_neighbor$label[image2_neighbor_labeled_indx] == image2_prediction_rf))

cat('predictive accuracy on the labeled pixels of image 3: \n')
print(mean(image3_neighbor$label[image3_neighbor_labeled_indx] == image3_prediction_rf))

# feature importance
cat('feature importances by decrease in GINI coefficient: \n')
print(sort(importance(rf_fit_neighbor) / sum(importance(rf_fit_neighbor)), decreasing = T))
```

# Lets try to smooth the features
We smooth the features using K-nearest-neighbors. We also add second order interaction terms. 
```{r}
# add second order interactions -- only add interactions for the non-engineered features
image1_poly_features <- add_interaction_terms(image1[, -c(1:6)])
image2_poly_features <- add_interaction_terms(image2[, -c(1:6)])

image1_poly_features <- cbind(image1[, 4:6], image1_poly_features)
image2_poly_features <- cbind(image2[, 4:6], image2_poly_features)

# smooth all features 
image1_poly_features_smoothed <- smooth_all_features(image1_poly_features, 
                                                     coord = image1[, 1:2])

image2_poly_features_smoothed <- smooth_all_features(image2_poly_features, 
                                                     coord = image2[, 1:2], 
                                                     k = 24)
n_features <- dim(image1_poly_features_smoothed)[2]

# create full dataframe
image1_data_features_smoothed <- cbind(as.factor(image1$label),
                                       image1_poly_features_smoothed)
colnames(image1_data_features_smoothed)[1] <- 'label'

image2_data_features_smoothed <- cbind(as.factor(image2$label),
                                       image2_poly_features_smoothed)
colnames(image2_data_features_smoothed)[1] <- 'label'

# this is so fucking annoying ... 
image1_data_features_smoothed_labeled <-
  image1_data_features_smoothed[image1_labeled_indx, ]
image1_data_features_smoothed_labeled$label <- droplevels(
  image1_data_features_smoothed_labeled$label)

image2_data_features_smoothed_labeled <-
  image2_data_features_smoothed[image2_labeled_indx, ]
image2_data_features_smoothed_labeled$label <- droplevels(
  image2_data_features_smoothed_labeled$label)

rf_fit_smo <- ranger(label ~ ., 
                data = image1_data_features_smoothed_labeled, 
                 num.trees = 500, 
                 probability = TRUE, 
                 importance = 'impurity')
```

Looking at predictive accuracy...
```{r , cache = TRUE}
image1_prediction_prob_rf_smo <- predict(rf_fit_smo,  
                          dat = image1_data_features_smoothed_labeled, 
                              type = 'response')
image1_prediction_rf_smo <-
  get_labels_from_prob(image1_prediction_prob_rf$predictions[, 2])


# predict on image2
image2_prediction_prob_rf_smo <- predict(rf_fit_smo, 
                              dat = image2_data_features_smoothed_labeled, 
                              type = 'response')

image2_prediction_rf_smo <-
  get_labels_from_prob(image2_prediction_prob_rf$predictions[, 2])

cat('Predictive accuracy on the label pixels of image 1: \n')
cat('note that image 1 was used to fit the RF: \n')
print(mean(image1$label[image1_labeled_indx] == image1_prediction_rf_smo))

cat('predictive accuracy on the labeled pixels of image 2: \n')
print(mean(image2$label[image2_labeled_indx] == image2_prediction_rf_smo))

```

examine feature importances
```{r}
cat('feature importances by decrease in GINI coefficient: \n')
print(sort(importance(rf_fit_smo) / sum(importance(rf_fit_orig)), decreasing = T))
```
