---
title: "Exploring the lasso"
author: "Bryan"
date: "10/27/2107"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r load_data, cache = TRUE, echo = FALSE}
library(knitr)
library(tidyverse)
library(glmnet)
# load data
image1 <- read.table('../image_data/image1.txt', header = F)
image2 <- read.table('../image_data/image2.txt', header = F)
image3 <- read.table('../image_data/image3.txt', header = F)

# add column labels
collabs <- c('y','x','label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
names(image1) <- collabs
names(image2) <- collabs
names(image3) <- collabs

```

```{r load_utils, cache = FALSE, echo = FALSE}
source('./lasso_lib.R')
source('./smoothing_fun.R')
```

# Fit using logistic regresssion with original 8 features (and pairwise interactions)
```{r get_features, cache = TRUE}
# get indices with labels
image1_labeled_indx <- which(image1$label != 0)
image2_labeled_indx <- which(image2$label != 0)
image3_labeled_indx <- which(image3$label != 0)

# get the original eight features
image1_features <- image1[, -c(1,2,3)]
image2_features <- image2[, -c(1,2,3)]
image3_features <- image3[, -c(1,2,3)]

image1_labels <- image1[, 3]
image2_labels <- image2[, 3]
image3_labels <- image3[, 3]

# add second order interactions
# image1_poly_features <- add_interaction_terms(image1[, -c(1,2,3)])
# image2_poly_features <- add_interaction_terms(image2[, -c(1,2,3)])
# image3_poly_features <- add_interaction_terms(image3[, -c(1,2,3)])

training_features <- as.matrix(rbind(image1_features[image1_labeled_indx, ],
                     image2_features[image2_labeled_indx, ]))
training_labels <- c(image1_labels[image1_labeled_indx], 
       image2_labels[image2_labeled_indx])

training_labels[training_labels == -1] <- 0

logist_fit <- glmnet(training_features, training_labels, family = 'binomial', 
                     lambda = 0)
```

```{r plot_logistic_results, cache = TRUE}
# get prediction probabilities
image1_prediction_prob_logistic <- predict(logist_fit, 
                              newx = as.matrix(image1_features),
                              type = 'response')
image2_prediction_prob_logistic <- predict(logist_fit, 
                              newx = as.matrix(image2_features),
                              type = 'response')
image3_prediction_prob_logistic <- predict(logist_fit, 
                              newx = as.matrix(image3_features),
                              type = 'response')

# round for plotting
image1_prediction_logistic <- get_labels_from_prob(image1_prediction_prob_logistic, 
                                                   thresh_lb = 0.5, 
                                                   thresh_ub = 0.5) 
image2_prediction_logistic <- get_labels_from_prob(image2_prediction_prob_logistic, 
                                                   thresh_lb = 0.5, 
                                                   thresh_ub = 0.5) 
image3_prediction_logistic <- get_labels_from_prob(image3_prediction_prob_logistic,
                                                   thresh_lb = 0.5, 
                                                   thresh_ub = 0.5) 

image1_orig <- 
  ggplot(image1) + geom_point(aes(x = x, y = y, 
                                  color = factor(label))) +
  scale_color_discrete(name = "Expert label")

image1_pred <- 
  ggplot(image1) + geom_point(aes(x = x, y = y, 
                                color = factor(image1_prediction_logistic))) +  
  scale_color_discrete(name = "Predicted label")

image2_orig <- 
  ggplot(image2) + geom_point(aes(x = x, y = y, 
                                  color = factor(label))) +
  scale_color_discrete(name = "Expert label")

image2_pred <- 
  ggplot(image2) + geom_point(aes(x = x, y = y, 
                                color = factor(image2_prediction_logistic))) +  
  scale_color_discrete(name = "Predicted label")

source('./multiplot.R')
multiplot(image1_orig, image2_orig, 
          image1_pred, image2_pred, cols=2)

```

Print coefficients: 
```{r}
print(logist_fit$beta)
```
examine accuracy: 
```{r get_logistic_accuracy, cache = TRUE}
cat('on the labeled pixels of image 1: \n')
print(mean(image1$label[image1_labeled_indx] ==
  image1_prediction_logistic[image1_labeled_indx]))

cat('on the labeled pixels of image2: \n')
print(mean(image2$label[image2_labeled_indx] ==
  image2_prediction_logistic[image2_labeled_indx]))

cat('on the labeled pixels of image3: \n')
print(mean(image3$label[image3_labeled_indx] ==
  image3_prediction_logistic[image3_labeled_indx]))
```


```{r logistic_smoothed}
# add smoothed features
image1_features_smo <- smooth_all_features(image1_features, 
                                                     coord = image1[, 1:2])
image2_features_smo <- smooth_all_features(image2_features, 
                                                     coord = image2[, 1:2])
image3_features_smo <- smooth_all_features(image3_features, 
                                                     coord = image3[, 1:2])

training_features_smo <- as.matrix(rbind(
  image1_features_smo[image1_labeled_indx, ],
  image2_features_smo[image2_labeled_indx, ]))

training_labels <- c(image1_labels[image1_labeled_indx], 
       image2_labels[image2_labeled_indx])

training_labels[training_labels == -1] <- 0

logist_fit_smo <- glmnet(training_features_smo, training_labels, family = 'binomial', 
                     lambda = 0)

```

look at prediction results
```{r}
# get prediction probabilities
image1_prediction_prob_logistic_smo <- predict(logist_fit_smo, 
                              newx = as.matrix(image1_features_smo),
                              type = 'response')
image2_prediction_prob_logistic_smo <- predict(logist_fit_smo, 
                              newx = as.matrix(image2_features_smo),
                              type = 'response')
image3_prediction_prob_logistic_smo <- predict(logist_fit_smo, 
                              newx = as.matrix(image3_features_smo),
                              type = 'response')

# round for plotting
image1_prediction_logistic_smo <-
  get_labels_from_prob(image1_prediction_prob_logistic_smo, 
                       thresh_lb = 0.5, 
                        thresh_ub = 0.5) 
image2_prediction_logistic_smo <-
  get_labels_from_prob(image2_prediction_prob_logistic_smo,
                       thresh_lb = 0.5, 
                        thresh_ub = 0.5) 

image3_prediction_logistic_smo <-
  get_labels_from_prob(image3_prediction_prob_logistic_smo, 
                       thresh_lb = 0.5, 
                        thresh_ub = 0.5) 

cat('on the labeled pixels of image 1: \n')
print(mean(image1$label[image1_labeled_indx] ==
  image1_prediction_logistic_smo[image1_labeled_indx]))

cat('on the labeled pixels of image2: \n')
print(mean(image2$label[image2_labeled_indx] ==
  image2_prediction_logistic_smo[image2_labeled_indx]))

cat('on the labeled pixels of image3: \n')
print(mean(image3$label[image3_labeled_indx] ==
  image3_prediction_logistic_smo[image3_labeled_indx]))

```

# Fit using LASSO 

### examining the lasso path
Here, we will explore cloud prediction using a lasso-logistic regression. First, we fit lasso using only the original 8 coefficients. Will we be able to recover the 3 engineered features? That is, will those 3 features (NDIR, CORR, SD) be found to be important? The lasso path is shown below: 
```{r lasso_orig, cache = TRUE, echo = FALSE}

# fit lasso and examine path
fit <- glmnet(training_features, training_labels, family = "binomial")
plot(fit, xvar = "lambda", label = TRUE)

```

### fitting the model
Here, we use 2-fold cross-validation (the two folds corresponding to the two images) to find $\lambda$ That is, we fit logistic-lasso using the first image for a range of $\lambda$. We then get the prediction accuracy (by simple mis-classifcation error) on the second image. We then reverse the roles of the two images in this procedure. The preditive accuracies are averaged, and plotted below as a function of $\lambda$. We then use the optimal $\lambda$ to fit on the entire dataset. 

Here, we fit using the orignal 8 as well as all pairwise interactions

```{r lasso_orig_pred, cache = TRUE, echo = FALSE}
fit_lasso <- get_fit_2fold_cv(as.matrix(image1_features[image1_labeled_indx, ]),
                             image1$label[image1_labeled_indx],
                             as.matrix(image2_features[image2_labeled_indx, ]), 
                             image2$label[image2_labeled_indx])
```

```{r lasso_orig_plots1, cache = TRUE, echo = FALSE}
ggplot() + geom_point(aes(x = fit_lasso$cv_results$lambda, 
                          y = fit_lasso$cv_results$accuracy)) + 
  ylab('classification accuracy') + xlab('lambda') +
  geom_vline(xintercept=fit_lasso$cv_results$lambda_best)

```

At this value of lambda, we print the coefficients: 
```{r, cahce = TRUE}
cat('coefficients: \n')
print(fit_lasso$fit$beta)
```


The result of this fit is shown below. Probabilities greater than 0.6 are labeled +1, probabilities less than 0.4 are labeled -1, and those in between are 0.  
```{r plot_lasso_predictions, cache = TRUE}
# get prediction probabilities
image1_prediction_prob_lasso_orig <- predict(fit_lasso$fit, 
                                             newx = as.matrix(image1_features),
                                             s = fit_lasso$cv_results$lambda_best,
                                             type = 'response')

image2_prediction_prob_lasso_orig <- predict(fit_lasso$fit, 
                                             newx = as.matrix(image2_features),
                                             s = fit_lasso$cv_results$lambda_best,
                                             type = 'response')

image3_prediction_prob_lasso_orig <- predict(fit_lasso$fit, 
                                             newx = as.matrix(image3_features),
                                             s = fit_lasso$cv_results$lambda_best,
                                             type = 'response')
# round
image1_prediction_lasso_orig <- get_labels_from_prob(image1_prediction_prob_lasso_orig)
image2_prediction_lasso_orig <- get_labels_from_prob(image2_prediction_prob_lasso_orig)
image3_prediction_lasso_orig <- get_labels_from_prob(image3_prediction_prob_lasso_orig)

```

The final accuracy of our predictions: 
```{r get_lasso_accuracy, cache = TRUE}
cat('on the labeled pixels of image 1: \n')
print(mean(image1$label[image1_labeled_indx] ==
  image1_prediction_lasso_orig[image1_labeled_indx]))

cat('on the labeled pixels of image2: \n')
print(mean(image2$label[image2_labeled_indx] ==
  image2_prediction_lasso_orig[image2_labeled_indx]))

cat('on the labeled pixels of image3: \n')
print(mean(image3$label[image3_labeled_indx] ==
  image3_prediction_lasso_orig[image3_labeled_indx]))

```

# Lets try to smooth the features
We smooth the features using K-nearest-neighbors. We also add interaction terms. 
```{r}
source('./smoothing_fun.R')
# smooth features
image1_features_smoothed <- smooth_all_features(image1_features, 
                                                     coord = image1[, 1:2])
image2_features_smoothed <- smooth_all_features(image2_features, 
                                                     coord = image2[, 1:2])
image3_features_smoothed <- smooth_all_features(image3_features, 
                                                     coord = image3[, 1:2])


fit_lasso_smooth <- get_fit_2fold_cv(
                            as.matrix(image1_features_smoothed[image1_labeled_indx, ]),
                             image1_labels[image1_labeled_indx],
                            as.matrix(image2_features_smoothed[image2_labeled_indx, ]), 
                             image2_labels[image2_labeled_indx])

ggplot() + geom_point(aes(x = fit_lasso_smooth$cv_results$lambda, 
                          y = fit_lasso_smooth$cv_results$accuracy)) + 
  ylab('classification accuracy') + xlab('lambda') +
  geom_vline(xintercept=fit_lasso_smooth$cv_results$lambda_best)

```
The coefficients in this case ... 
```{r, cahce = TRUE}
cat('coefficients: \n')
print(fit_lasso_smooth$fit$beta)
```

```{r plot_lasso_predictions, cache = TRUE}
# get prediction probabilities
image1_prediction_prob_lasso_smo <- predict(fit_lasso_smooth$fit, 
                            newx = as.matrix(image1_features_smoothed), 
                             s = fit_lasso_smooth$cv_results$lambda_best,
                             type = 'response')

image2_prediction_prob_lasso_smo <- predict(fit_lasso_smooth$fit, 
                            newx = as.matrix(image2_features_smoothed), 
                             s = fit_lasso_smooth$cv_results$lambda_best,
                             type = 'response')

image3_prediction_prob_lasso_smo <- predict(fit_lasso_smooth$fit, 
                            newx = as.matrix(image3_features_smoothed), 
                             s = fit_lasso_smooth$cv_results$lambda_best,
                             type = 'response')

# round for plotting
image1_prediction_lasso_smo <- get_labels_from_prob(image1_prediction_prob_lasso_smo)
image2_prediction_lasso_smo <- get_labels_from_prob(image2_prediction_prob_lasso_smo)
image3_prediction_lasso_smo <- get_labels_from_prob(image3_prediction_prob_lasso_smo)

image1_orig <- 
  ggplot(image1) + geom_point(aes(x = x, y = y, 
                                  color = factor(label))) +
  scale_color_discrete(name = "Expert label")

image1_pred <- 
  ggplot(image1) + geom_point(aes(x = x, y = y, 
                                color = factor(image1_prediction_lasso_smo))) +  
  scale_color_discrete(name = "Predicted label")

image2_orig <- 
  ggplot(image2) + geom_point(aes(x = x, y = y, 
                                  color = factor(label))) +
  scale_color_discrete(name = "Expert label")

image2_pred <- 
  ggplot(image2) + geom_point(aes(x = x, y = y, 
                                color = factor(image2_prediction_lasso_smo))) +  
  scale_color_discrete(name = "Predicted label")

source('./multiplot.R')
multiplot(image1_orig, image2_orig, 
          image1_pred, image2_pred, cols=2)


```

The final accuracy of our predictions: 
```{r get_lasso_accuracy, cache = TRUE}
cat('on the labeled pixels of image 1: \n')
print(mean(image1$label[image1_labeled_indx] ==
  image1_prediction_lasso_smo[image1_labeled_indx]))

cat('on the labeled pixels of image2: \n')
print(mean(image2$label[image2_labeled_indx] ==
  image2_prediction_lasso_smo[image2_labeled_indx]))


cat('on the labeled pixels of image3: \n')
print(mean(image3$label[image3_labeled_indx] ==
  image3_prediction_lasso_smo[image3_labeled_indx]))

# get indices with labels
image3_labeled_indx <- which(image3$label != 0)

# add second order interactions
image3_poly_features <- add_interaction_terms(image3[, -c(1,2,3)])
image3_poly_features_smoothed <- smooth_all_features(image3_poly_features, 
                                                     coord = image3[, 1:2])
image3_poly_features_smoothed <- as.matrix(image3_poly_features_smoothed)

image3_labels <- image3[, 3]
image3_prediction_prob_lasso_orig <- predict(fit_poly_smooth$fit, 
                              newx =image3_poly_features_smoothed, 
                             s = fit_poly_smooth$cv_results$lambda_best,
                             type = 'response')

image3_prediction_lasso_orig <-get_labels_from_prob(image3_prediction_prob_lasso_orig)

print('test accuracy on held out third image: ')
print(mean(image3$label[image3_labeled_indx] ==
  image3_prediction_lasso_orig[image3_labeled_indx]))

```

